\documentclass[letterpaper,hyper]{THG_RFC}

% Path to figures, plots
\graphicspath{{./pics/}{./plots/}}

%% TikZ
\usepackage{tikz}
\usetikzlibrary{patterns, shapes, positioning, decorations.pathreplacing,
backgrounds, fit}
\usepackage{gnuplot-lua-tikz}

% SI units
%\usepackage{textcomp}
\usepackage[binary-units]{siunitx}
\sisetup{per-mode = symbol}
\providecommand{\gbps}[1]{\SI{#1}{\giga\byte\per\second}}
\providecommand{\mbps}[1]{\SI{#1}{\mega\byte\per\second}}
\providecommand{\gb}[1]{\SI{#1}{\giga\byte}}
\providecommand{\mb}[1]{\SI{#1}{\mega\byte}}
\providecommand{\kb}[1]{\SI{#1}{\kilo\byte}}

% Break table
\usepackage{longtable}

% Code snippets
\usepackage{listings}
\def\lstsetc{\lstset{language=C,
  %numbers=left,
  xleftmargin=1.8em,
  frame=L,
  framexleftmargin=0.2em,
  rulecolor=\color{rulecolor},
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  showspaces=false, 
  showstringspaces=false,
  breaklines=true,
  basicstyle=\footnotesize\ttfamily,
  stringstyle=\itshape,
  commentstyle=\itshape\bfseries,
  morekeywords={main, size_t, malloc, free,
    hsize_t, hid_t, herr_t,
    H5X_class_t, H5X_type_t, H5X_info_t,
    H5Q_combine_op_t, H5Q_type_t, H5Q_match_op_t,
    H5Fcreate, H5Fclose, H5Fopen,
    H5Gcreate, H5Gopen, H5Gclose,
    H5Dcreate, H5Dopen, H5Dclose, H5Dwrite, H5Dread, H5Dquery, H5Dget_type,
    H5Dget_space,
    H5Pcreate, H5Pclose, H5Pset_index_plugin,
    H5Qcreate, H5Qclose, H5Qcombine, H5Qget_type, H5Qget_match_op,
    H5Qget_components, H5Qget_combine_op, H5Qencode, H5Qdecode,
    H5Qapply, H5Qapply_multi, H5Qapply_str, H5Qapply_dir,
    H5Rget_region,
    H5Screate_simple, H5Sclose, H5Sget_select_npoints,
    H5Tget_size,
    H5Xregister, H5Xunregister, H5Xcreate, H5Xremove, H5Xget_count, H5Xget_size,
    H5Xget_info
    }
  }
}

% Title, author, etc
\title{Data Analysis Extensions}
\author{Jerome Soumagne}
\author{Quincey Koziol}
\date{Feb 8, 2016}
\rfcversion{2014-07-17.v10}
\revision{Jul. 17, 2014}{Version 1 circulated for comment within The HDF Group.}
\revision{Aug. 20, 2014}{Version 2 circulated for comment within The HDF Group.}
\revision{Nov. 14, 2014}{Version 3 circulated for comment within The HDF Group.}
\revision{Feb. 25, 2015}{Version 4 edits and typo fixing.}
\revision{Jul. 24, 2015}{Version 5 edits to view object.}
\revision{Oct. 14, 2015}{Version 6 rework of query/view sections.}
\revision{Oct. 15, 2015}{Version 7 first draft of cross-container queries.}
\revision{Oct. 31, 2015}{Version 8 some corrections.}
\revision{Feb. 3, 2016}{Version 9 more text corrections.}
\revision{Feb. 8, 2016}{Version 10 add metadata indexing and usage examples.}

%% Start the document
\begin{document}

%% Title
\maketitle

%% Abstract
\begin{abstract}
Accessing, selecting and retrieving data from an HDF5 container can be a time
consuming process, particularly so when data is very large. To enable, ease and 
accelerate this process, we introduce a set of extensions to the HDF5 library
that enable efficient query, selection, and indexing of data. This document defines
the \texttt{H5Q} \textit{query} and \texttt{H5X} \textit{indexing} interfaces as
well as a \textit{view} structure that can be used to store and retrieve query results.
\end{abstract}

\input{analysis_ext-body}

\section*{Revision History}
\makerevisions

%% References
\bibliographystyle{abbrv}
\bibliography{analysis_ext}

\end{document}
